Don't Mind the Gap: Bridging Network-wide Objectives and Device-level Configurations

### 2016-06-16 paper/v2/paper-edited.tex
5.1

The front-end (FE) of \sysname simplifies operators' task of describing preferred paths. That simplicity, however, comes at the cost of compilation complexity. The compiler must efficiently compute the sets of paths represented by the intersection of preferences and topology, determine which ones can be honored under failures, and ensure policy compliance under all possible failure cases.

@@

  \caption{Regular Intermediate Language (RIL) syntax (left), and
           Propane language expansions (right).}
  \label{fig:rir-syntax}
  %\vspace{-1em}
\end{figure*}

@@

We handle these challenges by breaking compilation into multiple stages (see Figure~\ref{fig:pipeline}), and developing efficient algorithms for the translation between each stage. The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for well-formedness (\EG, never constraining traffic that does not enter the network), before being combined with the network topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines that the PGIR is safe, it translates it to an abstract BGP (ABGP) representation. ABGP can then be translated into various vendor-specific device
configurations as needed. The \sysname compiler currently generates
Quagga router configurations from ABGP.

\para{Syntax}
The \sysname FE is just a thin layer atop a core, regular-expression language (RIR) for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy consists of one or more constraints, each of which have a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expressions abbreviation $r^+$ to stand for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisments with a particular community value (i.e., an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The semantics of RIR is in terms of ranked paths. Each preference-based regular path constraint (of the form $r_1 \Prefer \dots \Prefer r_j$) maps to a set of concrete paths in the network that match one of $r_i$. We denote a network path as a string of abstract locations (routers or external ASes) of the form: $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$, that is the path is in the language of the regular expression, and $p$ is a topologically-valid path. We denote the length of a path $p$ as $\abs{p}$. A path $p$ will have a rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
where the rank is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the path length. Lower ranks indicate \emph{more} preferred paths. Traffic may be sent on any of the most preferred paths for each pair of starting and ending locations that appear in some specified path. There is an implicit lowest preference to drop traffic: $\emptyset$ when no other route is possible.

@@

\para{Propane FE to RIR}

The first step in compilation is to reduce the \sysname front-end language (FE) to the simpler, regular intermediate representation (RIR) from Figure~\ref{fig:rir-syntax}.
The main differences between the FE and RIR are: $i)$ FE allows the programmer to specify constraints using a series of (modular) definitions, and combine them later, $ii)$ FE provides high-level names that abstract sets of routes and groups of prefixes/neighbors, and $iii)$ FE allows the preference operator to be used more flexibly.

A key constraint when translating FE to RIR is to ensure that all specified routes are well-formed. In particular, each regular path constraint $r$ must satisfy $r \subseteq \Out^* \cdot \In^+ \cdot \Out^*$. This ensures that users only control traffic that goes through their network at some point, and that such traffic does not loop back multiple times through their network.

The translation from \sysname FE to RIR is based on a set of rewriting rules.
The first step merges separate constraints. It takes the cross product of per-prefix constraints, where logical conjunction ($a \AND b$) is replaced by intersection on regular constraints ($a \Intersect b$), logical disjunction is replaced by union, and logical negation ($\NOT a$) is replaced by path negation ($\Any \cap \NOT(a)$). The additional constraint $\Any$ ensures the routes are well-formed by restricting the paths to only those that go through the user's network.
%
For example, in the data center configuration from \S\ref{sec:propane}, combining the \CD{Locality} and \CD{Ownership} policies results in the following RIR:

@@

The next step rewrites the high-level constraints such as \textit{enter} according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing distributivity equivalences:
%
\[
\begin{array}{c}
  x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n) \\
  (y_1 \Prefer \dots \Prefer y_n) \odot x = (y_1 \odot x) \Prefer \dots \Prefer (y_n \odot x)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $x$ is a policy with a single preference. Preferences nested under a unary operator (i.e., \textit{star} or \textit{negation}) are flagged by the compiler as invalid policies.

>>>>

The examples above use what we call the front end (FE) of \sysname.
It simplifies operators' task of describing preferred paths, but that simplicity comes at the cost of compilation complexity. The compiler must efficiently compute the sets of paths represented by the intersection of preferences and topology, determine which ones can be honored under failures, and ensure policy compliance under all failure cases.

@@

  \caption{Regular Intermediate Representation (RIR) syntax (left), and
           Propane language expansions (right).}
  \label{fig:rir-syntax}
  %\vspace{-1em}
\end{figure*}

To handle these challenges, we decompose compilation into multiple stages, shown in  Figure~\ref{fig:pipeline}, and develop efficient algorithms for the translation between stages. The first stage of the pipeline involves simple rewriting rules and substitutions from the FE to the core Regular Intermediate Representation (RIR). Policies in RIR are checked for well-formedness (\EG, never constraining traffic that does not enter the network), before being combined with the network topology to obtain the Product Graph Intermediate Representation (PGIR). The PGIR is a data representation that compactly captures the flow of BGP announcements subject to the policy and topology restrictions. We develop efficient algorithms that operate over the PGIR to ensure policy compliance under failures, avoid BGP instability, and prevent aggregation-induced black holes. Once the compiler determines safety, it translates the PGIR to an abstract BGP (ABGP) representation. ABGP can then be translated into various vendor-specific device
configurations as needed.
%The \sysname compiler currently generates
% Quagga router configurations from ABGP.

@@

%\para{Syntax}
The \sysname FE is just a thin layer atop the RIR for describing preference-based path constraints.
Figure~\ref{fig:rir-syntax} shows the RIR syntax. A policy has one or more constraints, each of which has a test on the type of route, and a corresponding set of preferred regular paths. Regular paths are regular expressions where the base characters are abstract locations---either a router or an AS. Special \In{} and \Out{} symbols refer to any internal or external location respectively. In addition, $\Sigma$ refers to any symbol. We also use the standard regular expressions abbreviation $r^+$ for $r \cdot r^*$, a sequence of one or more occurrences of $r$. Predicates ($t$) consist of logical boolean connectives (and, or, not) as well as tests that match a particular prefix (or group of prefixes) and tests for route advertisments with a particular community value (i.e., an integer value associated with a path).
%% For example, the constraint
%% \[(\textit{prefix}=74.3.28.0/24) \Link \\
%% (as200 \cdot \In^+) \Prefer (as100 \cdot \In^+)
%% \]
%% describes a more-preferred set of paths for traffic announced by a prefix no less specific than \CD{74.125.28.0/24}, which starts at AS 200, before entering and staying inside the user's network to get to the destination, and a less-preferred set of paths that start at AS 100 and are otherwise the same. The plus operator $\In^+$ stands for $\In \cdot \In^*$, at least one internal hop. Tests over route types use standard boolean connectives, and can refer to both prefixes and route community values.

@@

The semantics of RIR is in terms of ranked paths. Each preference-based regular path constraint of the form $r_1 \Prefer \dots \Prefer r_j$ maps to a set of concrete paths in the network that match one of $r_i$. A network path is a string of abstract locations of the form $n_1 n_2 \dots n_k$. A regular expression $r$ matches path $p$, if $p \in \mathcal{L}(r)$, that is, the path is in the language of the regular expression and is topologically valid. 
%We denote the length of a path $p$ as $\abs{p}$.

A path $p$ has rank:
$$(\min_i \set{ p \in \mathcal{L}(r_i) }, \abs{p})$$
which is lexicographically ordered according to (1) the most preferred regular expression matched, and (2) as a tie breaker, the path length $\abs{p}$. Lower ranks indicate \emph{more} preferred paths. Traffic may be sent on any of the most preferred paths for each pair of starting and ending locations that appear in some specified path. There is an implicit lowest preference $\emptyset$ to drop traffic when no other route is possible.

@@

\para{From FE to RIR}

The first stage in \sysname compilation reduces the FE to the simpler RIR from Figure~\ref{fig:rir-syntax}.
The main differences between the FE and RIR are: $i)$ FE allows the programmer to specify constraints using a series of (modular) definitions, and combine them later, $ii)$ FE provides high-level names that abstract sets of routes and groups of prefixes/neighbors, and $iii)$ FE allows the preference operator to be used more flexibly.

A key constraint when translating FE to RIR is ensuring that all specified routes are well-formed. In particular, each regular path constraint $r$ must satisfy $r \subseteq \Out^* \cdot \In^+ \cdot \Out^*$. It ensures that users only control traffic that goes through their network at some point, and that such traffic does not loop back multiple times through their network.

The translation from FE to RIR is based on a set of rewriting rules.
The first step merges separate constraints. It takes the cross product of per-prefix constraints, where logical conjunction ($a \AND b$) is replaced by intersection on regular constraints ($a \Intersect b$), logical disjunction is replaced by union, and logical negation ($\NOT a$) is replaced by path negation ($\Any \cap \NOT(a)$). The additional constraint $\Any$ ensures the routes are well-formed by restricting the paths to only those that go through the user's network.
%
For example, in the data center configuration from \S\ref{sec:propane}, combining the \CD{Locality} and \CD{Ownership} policies results in the following RIR:

@@

The next step rewrites the high-level constraints such as $\Enter$ according to the equivalences in Figure~\ref{fig:rir-syntax}. Since preferences can only occur at the outermost level for an RIR expression, the final step is to ``lift" occurrences of the preference operator in each regular expression. For example, the regular expression $a \cdot (b \Prefer c) \cdot d$ is lifted to $(a \cdot b \cdot d) \Prefer (a \cdot c \cdot d)$ by distributing the preference over the sequence operator. In general, we employ the follwing distributivity equivalences:
%
\[
\begin{array}{c}
  x \odot (y_1 \Prefer \dots \Prefer y_n) = (x \odot y_1) \Prefer \dots \Prefer (x \odot y_n) \\
  (y_1 \Prefer \dots \Prefer y_n) \odot x = (y_1 \odot x) \Prefer \dots \Prefer (y_n \odot x)
\end{array}
\]
%
where $\odot$ stands for an arbitrary regular binary operator, and $x$ is a policy with a single preference. The compiler flags as invalid preferences nested under a unary operator (i.e., \textit{star} or \textit{negation}).
